---
title: "Benchmark"
author: "Jon Ahlinder"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/joah/OneDrive - Skogforsk/Documents/Projekt/Emelie/Nationell_samverkan/")
require(dplyr)
require(phyloseq)
require(ggplot2)
```

## Read benchmark data

```{r}
trait <- read.table(file = "Beetels Test Jon metadata.txt",sep="\t",header = TRUE)
data <- read.table(file = "Beetels Test Jon species matrix data.txt",sep="\t",header = TRUE)
metadata <- data %>% 
  select(c("Area_T","Treatment","Area","Type","Treatment.1","Area.1","Site"))
abund <- data %>%
  select(trait$Species)
rownames(abund) <- metadata$Site
rownames(metadata) <- metadata$Site
# 
# Create a phyloseq object
abund2 <- as.matrix(t(abund))
OTU <- otu_table(abund2,taxa_are_rows = TRUE)
sampledata <- sample_data(metadata)
# Kingdom	Animalia
# Division	Arthropoda
# Class	Insecta
# Order	Coleoptera
# Domain Phylum Class Order Family Genus Species
#nr <- nrow(trait)
#data.frame(Kingdom = rep("Animalia",nr),Division = rep("Arthropoda",nr),Class = rep("Insecta",nr),Order = rep("Coleoptera",nr),Family = trait$Family,Genus = )
#TAX <- tax_table(as.matrix())

physeq = phyloseq(OTU, sampledata)
```


## Plot data

```{r}
########### Alpha diversity #########
plot_richness(physeq, x="Area.1", color="Treatment",measures = c("Observed","Chao1","Shannon"))
########### Beta diversity #############
GP.ord <- ordinate(physeq, "NMDS", "bray")
p = plot_ordination(physeq, GP.ord, color="Area.1", shape="Treatment")
p = p + geom_point(size=3, alpha=0.75)
#p = p + scale_colour_brewer(type="qual", palette="Set1")
p + ggtitle("NMDS on Bray-Curtis distance")
```

# Test JSDM implementations

## sjSDM
https://github.com/TheoreticalEcology/s-jSDM
A new joint species distribution model for faster and more accurate inference of species associations from big community data: https://doi.org/10.1111/2041-210X.13687

```{r}
#install.packages("sjSDM")
require(sjSDM)
```

## HMSC

```{r}
#install.packages("Hmsc")
#install.packages("snow") # In the event that it is not installed with Hmsc
#install.packages("corrplot")
require(Hmsc)
```

## GLLVM

```{r}
require(gllvm)
```

## jSDM
jSDM R package 
For joint species distribution models 
https://ecology.ghislainv.fr/jSDM 
```{r}
install.packages("C:/Users/joah/Downloads/jSDM_0.1.2.zip", repos = NULL, type="source")
require(jSDM)
```

## JSDM4POD
https://github.com/molgor/JSDM4POD/
a Joint Species Distribution Model for Presence-Only Data. Original package in python calling PyStan. 
```{r}

```

